Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-01-06 manuscript/manuscript.Rmd
updated with intro beef up

Aggregate measures of individual plant performance, such as those typically collected as part of large-scale census efforts, offer an alternative to detailed demographic data for modeling populations. Such population-level data will never match the precision of individual-level data,

>>>>

Aggregate measures of individual plant performance, such as those typically collected as part of large-scale census efforts, offer an alternative to detailed demographic data for modeling populations [@Clark2004; @Freckleton2011]. Such population-level data will never match the precision of individual-level data, but it is more feasible to attain a broad coverage sample when collecting coarse-scale data. This presents a difficult trade-off: on the one hand, individual-level data leads to more reliable models; on the other hand, population-level data leads to models that will produce less precise predicitons but can be applied over greater spatial and temporal extents. If the goal of the model is to predict the consequences of climate change, then we argue that in the face of limited time and money population-level data are the best option. However, the performance of models based on population-level data have rarely been compared to that of models based on individual-level data.

To date, relatively few studies have tried to model populations based on data other than detailed individual-level data. An important exception is an effort by @Taylor2004 to model the population growth rate of an invasive species to investigate the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Building on this work, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 showed the application of such methods in a study on arable weeds. In particular, @Queenborough2011 provide emprical evidence that density-structured models are capable of reproducing population dynamics, even if some precision is lost when compared to fully continuous models. Thus, population models based on coarse, population-level data show promise for producing ecological forecasts at landscape and regional scales [@Queenborough2011]. However, none of these models included environmental covariates.

Basing population models on population- rather than individual-level data in a climate change context is hampered by the fact that it is individuals that respond to climate, not populations [@Clark2012]. This fact puts us in uneasy proximity to an "ecological fallacy" where one deduces inference on the individual from statistical inference on the group [@PIANTADOSI1988].

### 2015-01-06 manuscript/manuscript.tex
updated with intro beef up

Aggregate measures of individual plant performance, such as those
typically collected as part of large-scale census efforts, offer an
alternative to detailed demographic data for modeling populations. Such
population-level data will never match the precision of individual-level
data,

@@

@@

@@

>>>>

Aggregate measures of individual plant performance, such as those
typically collected as part of large-scale census efforts, offer an
alternative to detailed demographic data for modeling populations (Clark
and Bjørnstad 2004, Freckleton et al. 2011). Such population-level data
will never match the precision of individual-level data, but it is more
feasible to attain a broad coverage sample when collecting coarse-scale
data. This presents a difficult trade-off: on the one hand,
individual-level data leads to more reliable models; on the other hand,
population-level data leads to models that will produce less precise
predicitons but can be applied over greater spatial and temporal
extents. If the goal of the model is to predict the consequences of
climate change, then we argue that in the face of limited time and money
population-level data are the best option. However, the performance of
models based on population-level data have rarely been compared to that
of models based on individual-level data.

To date, relatively few studies have tried to model populations based on
data other than detailed individual-level data. An important exception
is an effort by Taylor and Hastings (2004) to model the population
growth rate of an invasive species to investigate the best strategies
for invasion control. They used a ``density-structured'' model where the
state variable is a discrete density state rather than a continuous
density measure. Building on this work, Freckleton et al. (2011) showed
that density-structured models compare well to continuous models in
theory, and Queenborough et al. (2011) showed the application of such
methods in a study on arable weeds. In particular, Queenborough et al.
(2011) provide emprical evidence that density-structured models are
capable of reproducing population dynamics, even if some precision is
lost when compared to fully continuous models. Thus, population models
based on coarse, population-level data show promise for producing
ecological forecasts at landscape and regional scales (Queenborough et
al. 2011). However, none of these models included environmental
covariates.

Basing population models on population- rather than individual-level
data in a climate change context is hampered by the fact that it is
individuals that respond to climate, not populations (Clark et al.
2012). This fact puts us in uneasy proximity to an ``ecological
fallacy'' where one deduces inference on the individual from statistical
inference on the group (Piantadosi et al. 1988).

@@

Clark, J. S., and O. N. Bjørnstad. 2004. Population time series: Process
variability, observation errors, missing values, lags, and hidden
states. Ecology 85:3140--3150.

Clark, J. S., D. M. Bell, M. Kwit, A. Stine, B. Vierra, and K. Zhu.
2012. Individual-scale inference to anticipate climate-change
vulnerability of biodiversity.

@@

Piantadosi, S., D. P. Byar, and S. B. Green. 1988. The Ecological
Fallacy. American Journal of Epidemiology 127:893--904.

@@

Taylor, C. M., and A. Hastings. 2004. Finding optimal control strategies
for invasive species: a density-structured model for Spartina
alterniflora. Journal of Applied Ecology 41:1049--1057.
